---
title: "Basic functions"
description: "This vignette shows the functions for querying the CAX tables."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

See the [Getting Started vignette](setup.html) for instructions on installing {rCAX}. Once you have installed {rCAX}, you can begin using the package by loading the library.

Load the library.
```{r setup}
library(rCAX)
```
Read the terms of use:
```
rcax_termsofuse()
```

## HLI queries

The main user function is `rcax_hli()` which returns HLI tables from Coordinated Assessments data eXchange with extra 
meta data such as NMFS_PopID as returned from the [CAP Fish HLIs Tabular Query](https://www.streamnet.org/data/hli/).

The  basic `rcax_hli()` functionality is shown with the NOSA HLI but the syntax is the same for all the HLI: 

* NOSA: Natural Origin Spawner Abundance
* SAR: Smolt to Adult Ratios
* PNI: Proportionate Natural Influence of supplementation hatcheries
* RperS: Recruits per Spawner
* JuvOut: Juvenile Outmigrants
* PreSmolt: Presmolt Abundance


## Show the columns for the NOSA tables

Get the table and show all the column names with definitions. Only first 10 are shown.
```{r}
head(rcax_hli("NOSA", type="colnames"))
```

## Get records for the NOSA HLI table

Here the columns returned are restricted by `cols`. The table is filtered with `flist` to be just the columns with `nmfs_popid` equal to 7. Note the `cols` argument is case insensitive, `NMFS_PopID` and `nmfs_popid` are the same, but the column names in the returned tables will all be lower case.
```{r}
tab <- rcax_hli("NOSA",
  flist = list(nmfs_popid = 7),
  cols=c("NMFS_PopID", "spawningyear", "tsaej", "nosaej"))
head(tab)
```

Return data for a single ESU. The ESU/DPS names must be exact and are case sensitive. Use `rCAX:::caxesu` to see the ESU/DPS names.
```{r}
tab <- rcax_hli("NOSA",
  flist = list(esu_dps = "Salmon, chum (Columbia River ESU)")
)
```
We can then plot:
```{r}
library(ggplot2)
# Convert tsaej to a number
tab$tsaej <- as.numeric(tab$tsaej)
# plot
ggplot(
  subset(tab, spawningyear>2000), 
  aes(x=spawningyear, y=log(tsaej), color=waterbody)) +
  geom_line(na.rm = TRUE) +
  ggtitle("log(total spawners)")
```

Keep in mind that not all ESU and DPS are in the CAX database for each HLI. Go to https://www.streamnet.org/data/hli/ to do a search to quickly see what is available in different HLI tables.

## Filtering

### Single value

The `flist` argument is a list with the columns and values you want to filter on. Here are some examples:

Filter based on one value.
```
f <- list(esu_dps = "Steelhead (Middle Columbia River DPS)")
f <-  list(popid = 7)
```


### Multiple values

Filter based on multiple values. In this case, data with popid 7, 8 or 9 are returned.
```
flist = list(popid = c(7,8,9))
```

Filter based on two columns. Here we getting the the summer run data for one ESU. The values in `flist` are not case sensitive so "Summer" will return both "Summer" and "summer".
```
flist = list(esu_dps = "Salmon, Chinook (Snake River spring/summer-run ESU)", run = c("Summer"))
```
Unfortunately there seems to be a server-side problem with passing in multiple values with multiple columns. This works
```
flist = list(run = c("Summer", "Spring"))
```
But this throws an error.
```
flist = list(esu_dps = "Salmon, Chinook (Snake River spring/summer-run ESU)", run = c("Summer", "Spring"))
```

## Change the number of records returned

The default maximum number of records is 1000. You can increase (or decrease) this by passing in the limit query parameter.

```{r}
tab <- rcax_hli("NOSA",
  qlist = list(limit=1),
  cols=c("popid", "spawningyear", "tsaej"))
tab
```

Increase the limit to 2000 to ensure all the data are returned. Not run.
```{r eval=FALSE}
tab <- rcax_hli("NOSA",
  flist = list(esu_dps="Salmon, Chinook (Snake River spring/summer-run ESU)")
  qlist = list(limit=2000),
  cols=c("popid", "spawningyear", "tsaej"))
tab
```

## Show the available tables

Only the name and id columns are shown.
```{r}
tab <- rcax_tables(cols=c("name", "id"))
head(tab)
```

## Show internal data

These are internal data sets. Access with `rCAX:::`

* `caxesu` The ESU and DPS names, which appear in the `esu_dps` column in tables.
* `caxpops` The Populations table with all the population metadata, like NMFS_PopID and MPG.
* `caxsuperpops` The SuperPopulations table.

```{r}
rCAX:::caxesu[1:5]
```

```{r}
colnames(rCAX:::caxpops)
```

