---
title: "Getting Started"
description: "How to install and setup your API key"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an R package. Open R (e.g. in RStudio or a terminal), and issue the following commands at the command line.

## Installation

To install from GitHub, you will need the {remotes} package.

```{r eval=FALSE}
install.packages("remotes") # needed for the next line
```

Then you can install the latest release:
```{r eval=FALSE}
remotes:::install_github("nwfsc-math-bio/rCAX@*release")
```

Note, you only need to install packages once. After that you load them with a `library(pkgname)` call.

## Set up your API key

Next you need to get a pull api key from StreamNet (contact them). Once you have a key, you need to install the {usethis} package.

```{r eval=FALSE}
install.packages("usethis") # needed for the next line
```

Run this code to find or create your `.Renviron` file. It will tell you where the `.Renviron` file is.
```{r eval=FALSE}
usethis::edit_r_environ()
```
Open that file and paste in `CAX_KEY = "whateveryourkeyis"` into the file and save. 

Now restart R (Session > Restart R in RStudio) and your API will be visible to {rCAX}. You are now ready to use {rCAX}. Go to the [Basic functions vignette](articles/basics.html) to get started and see examples.

