---
title: "Getting Started"
description: "How to install and load rCAX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an R package. Open R (e.g. in RStudio or a terminal), and issue the following commands at the command line.

## Installation

To install from GitHub, you will need the {remotes} package.

```{r eval=FALSE}
install.packages("remotes") # needed for the next line
```

Then you can install the latest release:
```{r eval=FALSE}
remotes:::install_github("nwfsc-cb/rCAX@*release")
```

Note, you only need to install packages once. 

After installing, you load load the package with a `library(pkgname)` call. 

```{r}
library(rCAX)
```
To check if you have the latest rCAX use
```{r}
rcax_version()
```
If your version is out of date, update with another call to `install_github()` (above).

## API key

{rCAX} includes a "read-only" API key. Thus you
do not need an API key to use {rCAX}. If you need to use your own key, see vignette on the [API GET function](api_get.html).
